---
title: "doughtr"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

droughtr is an R for package that enables drought monitoring and drought forecasting. It computes the Standardized Precipitation Index (SPI) and the Non-Standardized Precipitation Index (NSPI) using General Additive Models for Location Scale and Shape (GAMLSS). It enables the generation of drought forecasts for univariate time series and deals with the bias introduced to drought forecasting applications when SPI and NSPI are calculated incorrectly during out-of-sample (OOS) validation.

## Installation

```{r, eval=FALSE}
# Install the development version on Github
devtools::install_github("mammask/droughtr")
```

## Usage

```{r, eval=TRUE}
# Load droughtr library
library(droughtr)

# Generate synthetic monthly rainfall data using the Gamma distribution
rain = dummyrainfall(startYear = 1950, endYear = 2010)

# Compute the non-stationary standardized precipitation index (NSPI) for scale 12 using GAMLSS
drought = computenspi(monthlyRainfall = rain, stationary = FALSE, spiScale = 12)

# Plot NSPI
plot(drought)
```

[]("\<README_files/figure-gfm/unnamed-chunk-2-1.png\>\#\<10\>x\<10\>")
