---
title: "droughtr"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

droughtr is an R for package that enables drought monitoring and drought forecasting. It computes the Standardized Precipitation Index (SPI) and the Non-Standardized Precipitation Index (NSPI) using General Additive Models for Location Scale and Shape (GAMLSS). It enables the generation of drought forecasts for univariate time series and deals with the bias introduced to drought forecasting applications when SPI and NSPI are calculated incorrectly during out-of-sample (OOS) validation.

```{r, echo = FALSE}
knitr::opts_chunk$set(
  fig.path = "README_figs/README-"
)
```

## Installation

```{r, eval=FALSE}
# Install the development version on Github
devtools::install_github("mammask/droughtr")
```

## Usage

### Generate SPI and NSPI

```{r, eval=TRUE, fig.height=3, fig.width=5}
# Load droughtr library
library(droughtr)

# Generate synthetic monthly rainfall data using the Gamma distribution
rain = dummyrainfall(startYear = 1950, endYear = 2010)

# Compute the non-stationary standardized precipitation index (NSPI) for scale 12 using GAMLSS
drought = computenspi(monthlyRainfall = rain, stationary = FALSE, spiScale = 12)

# Plot NSPI
plot(drought)
```

### Reduced-Bias forecasting framework

#### Data Split

Split the rainfall series into training validation and test set.

```{r, eval=TRUE}
rain = oossplit(x = rain, trainratio = 0.6, validationratio = 0.2, testratio = 0.2)
print(rain)
```

#### Bias measurement

When the Standardized Precipitation Index is calculated as part of a forecasting task it can potentially introduce biases in the training data. This is mainly because in many cases the index is computed using the entire data, prior to model validation, and this violates some of the fundamental principles of time series forecasting theory.

In this section, we compute the amount of bias introduced to the training set by measuring the number of miss-classifications in the training data. Two computational approaches are presented: 1) SPI is computed using the training data only; we call this as "Bias Corrected" computation and 2) SPI is computed using the entire data; we call this as "Bias Induced" computation. 

We measure bias by computing the number of miss-classifications in the training set due to the incorrect computation of the index. We also measure the number of records impacted and share a plot of the two computational approaches.

```{r, eval=TRUE}
# Generate synthetic monthly rainfall data using the Gamma distribution
rain = dummyrainfall(startYear = 1950, endYear = 2010)

# Compute bias
bias = measurebias(x = rain, trainratio = 0.6, validationratio = 0.2, testratio = 0.2, stationary = TRUE, spiscale = 12)
bias
```
